---
title: "Functionalizing Rasters from NEON"
author: "Amy Hudson"
date: "June 21, 2016"
output: html_document
---

## Objectives

Start by putting up objectives/tasks that students will be working though:
1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)

```{r import-raster}
library(raster)

teakCHM <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")

```

1. For the CHM, set values == 0 to NA (not trees)

```{r subset-raster}
teakCHM[teakCHM == 0] <- NA

hist(teakCHM)
abline(v=2, col= "red")
abline(v=6, col="red")
```

1. Classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize histogram/density and plot vertical cutoff lines.

```{r classify-raster}

height.class.m <- c(0, 6, 1,
                    6, 30, 2,
                    30, 50, 3,
                    50, 100, 4)

rcl.m <- matrix(height.class.m,
                ncol = 3,
                byrow = T)
#reclassify raster 
reclassified.chm <- reclassify(teakCHM,rcl.m)
hist(heightCHM)
density(teakCHM,
        main="Canopy Heights at Teakettle with bin cuttoffs in red",
        xlab = "Height (m)")
abline(v=rcl.m[,2],col="red")

#plot to pdf file
pdf(file= "TEAK_CHM_density_with_breaks.pdf", width = 6, height = 7)
density(teakCHM,
        main="Canopy Heights at Teakettle with bin cuttoffs in red",
        xlab = "Height (m)")
abline(v=rcl.m[,2],col="red", xpd=F)
dev.off()

# plot reclassified chm raster
par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 4.5))
plot(reclassified.chm,
     col=c("yellow","orange","blue","green"), 
     main="Canopy height classes at Lower Teakettle",
     legend=F)
# allow legend to plot outside of bounds
par(xpd=TRUE)
# create the legend x position
leg.x <- par()$usr[2] + 20
#legend y position
leg.y <- par()$usr[4] + 50 - abs(par()$usr[3]-par()$usr[4])/2
#add legend

legend(leg.x, leg.y,
       legend = c("0-6 m", "6-30 m", "30-50 m", "50-100 m"),  # make sure the order matches the colors, next
       fill = c("yellow", "orange","blue", "green"),
       bty="n") # turn off border

#plot to pdf file
pdf(file= "TEAK_CHM_classes.pdf", width = 4, height = 7)
par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 4.5))
plot(reclassified.chm,
     col=c("yellow","orange","blue","green"), 
     main="Canopy height classes at Lower Teakettle",
     legend=F)
par(xpd=TRUE)
leg.x <- par()$usr[2] + 20
leg.y <- par()$usr[4] + 50 - abs(par()$usr[3]-par()$usr[4])/2

legend(leg.x, leg.y,
       legend = c("0-6 m", "6-30 m", "30-50 m", "50-100 m"),
       fill = c("yellow", "orange","blue", "green"),
       bty="n")
dev.off()

#save reclassified raster to geotiff

writeRaster(reclassified.chm,
            filename = "reclassified_CHM_TEAK_GeoTIFF.tif",
            format = "GTiff",
            options = "COMPRESS=LZW",
            overwrite = T,
            NAflag = -9999)
```

1. Plot the classified raster, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)
1. Export the plot figure to a pdf – publishable
1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.



